Learning and comparing functional connectomes between populations

### 2012-12-30 paper.tex
added draft predictive modeling

@@

>>>>

1. importance of interpretation!
2. the need for feature extraction
3. the need for feature selection
4. setting parameters.
\section{Predictive Modeling}
craddock 2009, Dosenbach 2010, lord 2011, deshpande,

@@

Predictive modeling has been an active area of research in functional
neuroimaging since the early 90s [] that has more recently been adapted to
comparing connectomes [].  A predictive model defines a mathematical
relationship between a set of data features and a categorical or continuous
variable of interest (target) that is capable of decoding or \emph{predicting}
the target variable from the data.  In the context of comparing connectomes,
features correspond to bivariate measures of connectivity or other region
specifc graph statistics and the target is a phenotypic variable (e.g. age, sex,
disease state). The accuracy of a model for predicting the target variable
provides a natural means for evaluating the utility of the model as a biomarker
[] and for measuring the amount of information present in the connectome about
the target variable [Kjems, Kriegeskorte]. Prediction accuracy when combined
with a measure of model reproducibilty provides a means to evaluate experimental
tradeoffs for data acquisition, preprocessing, and analysis [NPAIRS]. Predictive
modeling techniques have been successfully applied to identify connectome-based
biomarkers of Alzheimer's disease [92], major depression [93], schizophrenia
[94], autism [47], ADHD [], and aging [].  Recent work has illustrated the
utility of predictive modeling for deriving connectivity models at the
individual level [97,].

Nearly every analytical method can be appplied in a predictive modelling framework, but
multivariate classification and regression have been the most commonly used
(referred to as MVPA; multivariate prediction analysis).  Multivariate methods
are attractive in connectomics because they are sensitive to dependencies
between features and avoid the need to correct for multiple comparisions since
the significance of an entire pattern is evaluated using a single statistical
test. Predictive modeling is typically performed in a supervised learning
framework where a target is available for each observation of the features,
although unsupervised and semi-supervised methods can also be used.

Predictive models are validated using one or more iterations of cross
validation, which involves training and testing phases.  A cross validation
cycle begins by seperating the available data into two independent subsets:
training data and testing data.  Training involves learning a predictive model
from the training data. During testing the model is applied to the testing data
to predict corresponding values of the target variable.  Prediction accuracy is
then assessed by comparing these predictions to the observed values of the
target variable. In order to make

A general application of predictive modeling can be broken down into feature
extraction, model selection, feature selection, and model validation, each of
which may contain one or more iteration of training and testing. Feature
extraction

Model training can be further broken down into feature extraction, model
selection and feature selection and testing can be performed for model
optimziation and

The application of predictive modeling to functional neuroimaging data has been
an active area of research since the early 90s[] and more recently for comparing
connectomes[]. Nearly every analytical method can be appplied in a predictive
modelling framework, but multivariate classification and regression have been
the most commonly used (referred to as MVPA; multivariate prediction analysis).
Multivariate methods are attractive in connectomics because they do not require
any assumptions regarding independence between features and avoid the need to
correct for multiple comparisions since the significance of an entire pattern is
evaluated using a single statistical test. Support vector machines
(classification and regression) are popular for MVPA applications owing to their
proven track record for acheiving good prediction accuracy in high dimensional
datasets (many features) with few observations.

Cross validation

MVPA algorithms typically require the specification of several parameters, the
values of which will impact the quality of the results. These parameters may be
prespecified based on domain specific knowledge, or requirements, determined
using an analytical approach, or optimized using cross validation. In latter
case a nest cross-validation scheme is required to avoid biasing resultant
measures of prediction accuracy. it is necessary to perform a nested
cross-validation These parameters may be set by hand in the case where

Determining the subset of features most relevant to the model (feature
selection) is essential for interpreting the models, and is the subject of
ongoing research. In cases were model weights can be directly mapped to their
corresponding feature, there is typically no statistical framework for
thresholding these weights, and meaningless features are not always set to zero
(this is particularly the case for L2 regularized algorithms such as SVM and
ridge regression). One feature selection strategy is to filter out features
based on their statistical relationship with the variable of interest using a
univariate statistical criterion such as t-test, Pearson's correlation, or
mutual information.  A drawback of this approach is that since the filters are
typically insensitive to multivaratiate information, they may exclude features
that would be important to the model.  Another approach is to treat feature
selection as a model selection problem in which case the features are
iteratively pruned and a subset is chosen based on prediction performance. A
third approach is to use sparse modeling methods such as L1 SVM or LASSO that
force meaningless weights to be zero.  Although ideal, these methods may have
limits on the number of non-zero weights used in the model or the exclusion of
highly correlated features from the model.  Contrary to previous reports [Guyon
2002, Craddock 2009] no matter which feature selection strategy is employed, it
must be performed withen cross-validation to avoid biasing estimates of
prediction accuracy [later Guyon].

The interpretation of classifier weights for selected features, when available,
is complicated and requires insight into the mechanisms underlying the modeling
approach. If features are appropriately standardized prior to training, the
magnitude of weights can be interpreted to reflect the relative importance of
the feature to the model, but interpreting how the feature differs between
classes or relates to a continuous variable is more complicated, particularly
given the multivariate nature of their involvement.  For classification methods
such as Fisher's linear descriminate analysis (LDA),which maximally seperates
class means given their standard deviation, feature weights are representative
of the differences between class means. For support vector classification,
classifier weights are derived from observations that are the most similar
between classes, and hence represents the border between classes.  It is perhaps
most reasonable to adapt a conservative interpretation in which MVPA is used to
identify the most relevant features that are later tested in followon
experiments better suited to elucidating their relationship to the variable of
interest.
